---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(factoextra)
library(ggfortify)
library(ggplot2)
library(dplyr)
```

```{r}
data <- read.csv("https://raw.githubusercontent.com/reisanar/datasets/master/macmenu.csv")
```

```{r}
data #make sure data loaded
```


```{r}
summary(data)
```

```{r}
#summary(menu_items_stats[1:3])
```



```{r}
dailyCals <- 2000
caloriePercents <- data %>%
  mutate(DailyCaloriePercent = (Calories / dailyCals) * 100) %>% 
  select(Item, DailyCaloriePercent) %>% 
  arrange(desc(DailyCaloriePercent))
```


```{r}
#this chunk of code is to make a data frame with only numeric values
x <- data[ , 4:24]
```


```{r}
stats <- data_frame(
  var_name = colnames(x),
  Avg = map_dbl(x, mean),
  Min = map_dbl(x, min) ,
  Max = map_dbl(x, max),
)
```


```{r}
breakfast <- data %>% 
    filter(Category == "Breakfast")
beefpork <- data %>% 
    filter(Category == "Beef & Pork")
chickenfish <- data %>% 
    filter(Category == "Chicken & Fish")
salads <- data %>% 
    filter(Category == "Salads")
snacksides <- data %>% 
    filter(Category == "Snacks & Sides")
desserts <- data %>% 
    filter(Category == "Desserts")
beverages <- data %>% 
  filter(Category == "Beverages") 
coffeeTea <- data %>% 
  filter(Category == "Coffee & Tea")
smoothieShakes <- data %>% 
  filter(Category == "Smoothies & Shakes")
```

```{r}
ggplot(data = data, aes(x = Category, fill = Category)) + 
  geom_bar(stat = "count", show.legend = FALSE) +
  coord_flip() + 
  labs(title = "Total Number of Products by Category", x = "Category", y = "Number of Items")
```

```{r}
ggplot(data = data, aes(x = Category, y = mean(Calories), fill = Category)) + 
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  labs(title = "Total Number of Products by Category", x = "Category", y = "Number of Items")
```



```{r}
ggplot(breakfast, aes(x = reorder(Item, Calories), y = Calories, fill = Calories)) +
  geom_bar(stat = "identity") + 
  coord_flip() +
  labs(title = "Breakfasts by Calories", x = "") +
  theme(text = element_text(size = 9.2))
```

```{r}
ggplot(beefpork, aes(x = reorder(Item, Calories), y = Calories, fill = Calories)) +
  geom_bar(stat = "identity") + 
  coord_flip() +
  labs(title = "Beef and Pork by Calories", x = "")
```

```{r}
ggplot(chickenfish, aes(x = reorder(Item, Calories), y = Calories, fill = Calories)) +
  geom_bar(stat = "identity") + 
  coord_flip() +
  labs(title = "Chicken and Fish by Calories", x = "")
```

```{r}
ggplot(desserts, aes(x = reorder(Item, Calories), y = Calories, fill = Calories)) +
  geom_bar(stat = "identity") + 
  coord_flip() + 
  labs(title = "Desserts by Calories", x = "")
```

```{r}
ggplot(snacksides, aes(x = reorder(Item, Calories), y = Calories, fill = Calories)) +
  geom_bar(stat = "identity") + 
  coord_flip() + 
  labs(title = "Snacks and Sides by Calories", x = "")
```

```{r}
ggplot(salads, aes(x = reorder(Item, Calories), y = Calories, fill = Calories)) +
  geom_bar(stat = "identity") + 
  coord_flip() + 
  labs(title = "Salads by Calories", x = "")
```

```{r}
ggplot(beverages, aes(x = reorder(Item, Calories), y = Calories, fill = Calories)) +
  geom_bar(stat = "identity") + 
  coord_flip() + 
  labs(title = "Beverages by Calories", x = "")
```

```{r}
coffeeTeaHigh <- coffeeTea %>% 
  arrange(desc(Calories)) %>% 
  head(47)

  ggplot(coffeeTeaHigh, aes(x = reorder(Item, Calories), y = Calories, fill = Calories)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size=8.5)) +
  coord_flip() +
  labs(title = "High Calorie Coffees and Teas",
       x = "")
```

```{r}
coffeeTeaLow <- coffeeTea %>% 
  arrange(desc(Calories)) %>% 
  tail(48)

  ggplot(coffeeTeaLow, aes(x = reorder(Item, Calories), y = Calories, fill = Calories)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size=9.1)) +
  coord_flip() +
  labs(title = "Low Calorie Coffees and Teas",
       x = "")
```

```{r}
  ggplot(smoothieShakes, aes(x = reorder(Item, Calories), y = Calories, fill = Calories)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size=9.5)) +
  coord_flip() +
  labs(title = "Smoothies and Shakes by Calories",
       x = "")
```

```{r}
#select only continuous attributes
menu_items_stats <- data[-c(1,2,3)]

#get item names
menu_items_names <- data[2]
```

From the principal component analysis below, we see that the first two principal components explain ~98% of the variation in the collection of 260 data points.
```{r}
#Principal Component Analysis
pca_items_stats <- prcomp(menu_items_stats)
summary(pca_items_stats)
```

Check the loadings for these 3 components:
```{r}
#check new loadings
pca_items_stats$rot[ , 1:2]
```

Visualization of PCA contributions
```{r}
fviz_screeplot(pca_items_stats)
```


PC1 largest loadings come from the sodium content and calories of the menu items.
PC2 largest loadings come from the calories and calories from fat. 

Bi-Plot visualization:
```{r}
autoplot(pca_items_stats, data = data, colour = "Category", loadings.colour = 'blue',
         loadings = TRUE, loadings.label = TRUE,
         loadings.label.size = 4, scale = 0)
```

```{r}
autoplot(pca_items_stats,data = data, colour = "Category", loadings.colour = 'blue',
         loadings = TRUE, loadings.label = FALSE)
```

Fviz biplot
```{r}
fviz_pca(pca_items_stats, geom = "point", repel = TRUE)
```

Contributions to the first principal component
```{r}
fviz_contrib(pca_items_stats, choice = "var", axes = 1) +
  labs(title = "Contribution of Variables to 1st Principal Component")
```
Contributions to the second principal component
```{r}
fviz_contrib(pca_items_stats, choice = "var", axes = 2) + 
  labs(title = "Contribution of Variables to 2nd Principal Component")
```

From the above principal component analysis, we can determine that calories, calories from fat, and sodium content appear to be the largest contributers to the principal components. 

Focusing on these attributes, let's try to generate clusters in our dataset.
```{r}
#reduce dataset to attributes identified from PCA
reduced_menu_items <- menu_items_stats[-c(3:9, 11:21)]

#scale data
reduced_menu_scaled <- scale(reduced_menu_items)

#determine optimal number of clusters
fviz_nbclust(reduced_menu_scaled, kmeans, method = "wss") + 
  labs(title = "Optimal Number of Clusters")
```

The visualization above suggests that we will capture most of the data by using 3 or 4 clusters.
```{r}
#kmeans clustering
df_kmeans <- kmeans(reduced_menu_scaled, centers = 4)

str(df_kmeans)
```

```{r}
#convert assignment to factor variable
km_results <- as.factor(df_kmeans$cluster)

#visualize Calories vs calories from fat
ggplot(reduced_menu_items, aes(x = Calories, y = Calories.from.Fat, color = km_results)) + 
      geom_point(show.legend = FALSE) + 
  labs(title = "Calories vs. Calories From Fat", y = "Calories From Fat")
```
```{r}
#visualize calories vs sodium
ggplot(reduced_menu_items, aes(x = Calories, y = Sodium, color = km_results)) + 
      geom_point(show.legend = FALSE) +
  labs(title = "Calories vs. Sodium")
```





